Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-03-14 introduction.tex
updates and finished seismic catalog section

However useful, most previous studies focus on particular locations or are based on specific categorization of seismogenic faults, which are often subjectively interpreted and limited to the availability data. Here, we use the smoothed seismicity method introduced by \citet{Frankel1995}. This approach has been used with success in other regions \citep[e.g.][]{Boyd2008, Kalkan2009, Moschetti2014}. Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that probabilistic seismic hazard analysis based on fault systems can be more realistic and accurate, by choosing the smoothed seismicity method we seek to reduce the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh; but for completeness, we also consider the partial contribution of the areas of Zagros and the central-east seismic zones within our domain of interest (see Figure \ref{fig:Iran}). In addition, for comparison purposes, we obtain results for the the combination of all seismic zones as a uniform region.

In the following sections we describe in greater detail our region of interest, review the associated tectonics and seismicity, and provide additional information about the approach followed to estimate seismic hazard for the region. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness, the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region \citep{Zafarani2014, Kalkan2004}, and the estimation of the Gutenberg-Richter seismicity parameters, which are necessary for the smoothed seismicity method. Our analysis includes seismic hazard curves and maps for peak ground acceleration considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions. We consider two hazard models based on the assumption that damage occurs above an earthquake magnitude $M_w$ 4.5 for informal construction buildings, and above $M_w$ 5.0 for structures built in accordance to code regulations. The results of these two models are then combined 
% and 5.0 for unreinforced and reinforced structures, respectively; We later combine these results 
into a single analysis based on the composition of the structural building inventory of the country. We compare our results to equivalent estimations (hazard curves) available in the literature for selected locations and conclude with a discussion in light of other previous hazard studies.

>>>>

However useful, most previous studies focus on particular locations or are based on specific categorization of seismogenic faults, which are often subjectively interpreted and limited to the availability data. Here, we use the smoothed seismicity method introduced by \citet{Frankel1995}. This approach has been used with success in other regions \citep[e.g.][]{Boyd2008, Kalkan2009, Moschetti2014}. Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that probabilistic seismic hazard analysis based on fault systems can be more realistic and accurate, by choosing the smoothed seismicity method we seek to reduce the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh; but for completeness, we also consider the partial contribution of the areas of Zagros and the central-east seismic zones within our domain of interest (see Fig.~\ref{fig:iran}). In addition, for comparison purposes, we obtain results for the the combination of all seismic zones as a uniform region.

In the following sections we describe in greater detail our region of interest, review the associated tectonics and seismicity, and provide additional information about the approach followed to estimate seismic hazard for the region. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness, the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region \citep{Zafarani2014, Kalkan2004}, and the estimation of the Gutenberg-Richter seismicity parameters, which are necessary for the smoothed seismicity method. Our analysis includes seismic hazard curves and maps for peak ground acceleration considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions. We consider two hazard models based on the assumption that damage occurs above an earthquake magnitude $M_w$ 4.5 for informal construction buildings, and above $M_w$ 5.0 for structures built in accordance to code regulations. The results of these two models are then combined into a single analysis based on the composition of the structural building inventory of the country. We compare our results to equivalent estimations (hazard curves) available in the literature for selected locations and conclude with a discussion in light of other previous hazard studies.

### 2016-03-14 seismic-catalog.tex
updates and finished seismic catalog section

\begin{figure*}[ht!] 
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-03}
	\caption{Seismic catalog. (a) Historical seismicity prior to 1900. (b) Instrumental seismicity. Size of symbols is proportional to the magnitude of the events. In the case of the historical seismicity catalog, the color fill indicates the year of the event.}
	\label{fig:selected}
\end{figure*}

@@

In this study we consider earthquakes with $M_w \geq 3$...

 % Earthquakes with magnitude 3 are not considered as a structural threat, however, the epicenter of earthquakes with magnitude $M_w 3$ are a  susceptible location for future bigger earthquakes.  According to our preliminary data processing based on IIEES data, Iranian catalog for earthquake with magnitude greater and equal 3, is complete from 2005. In order to make sure about the completeness of the catalog, we use IIEES data from 2000. We used the EMME project data provided by \citet{Zare2014} up to 2000. Fig.~\ref{fig:historical} represents the historical data (pre 1900) in the study region.

>>>>

\begin{figure*}[t] 
	\centering
	\includegraphics[width=0.9\textwidth]{figures/pdf/figure-03}
	\caption{Location of events considered in the compiled seismic catalog. (a) Historical seismicity prior to 1900 for $M_w \geq 4$ \citep[after][]{Zare2014}. (b) Instrumental seismicity up until December 2015 for $M_w \geq 3$ (IIEES) and $M_w \geq 4$ \citep[after][]{Zare2014}. Size of symbols is proportional to the magnitude of the events. In the case of the historical seismicity catalog, the color fill indicates the year of the event. The borderlines of the seismic zones shown in Fig.~\ref{fig:regional} are shown in the background in white, along with the light gray shaded relief.}
	\label{fig:catalog}
\end{figure*}

@@

In this study we adopt the catalog compiled by \citet{Zare2014} for both historical and instrumental earthquakes. This catalog includes events of magnitude $M_w \geq 4$ from 550 A.d.~to December 1999. We complemented it with events $M \geq 3$ occurred between January 2000 and December 2015 obtained from the IIEES database. Our analysis of this dataset indicated that the catalog satisfies completeness from 2005 onward. While we recognize that earthquakes in the range $3 \leq M_w \leq 4$ are unlikely to cause structural damage, we include them as a means to feed the model with information about epicentral locations with potential for future larger magnitude earthquakes. 
% The result of our compilation is shown in Fig.~\ref{fig:catalog}, which includes maps for historical and instrumental seismicity. 
% The following sections provide additional details about magnitude conversion and de\-clustering of the events in our combined catalog.

The portion of the catalog obtained from IIEES for $M \geq 3$ earthquakes between 2000 and 2015 contained event reports that use different magnitude scales. We converted all events with magnitudes in $M_L$, $M_s$, and $mb$ scales to $M_w$ using the same conversion relationships used by \citet{Zare2014}, which allowed us to preserve compatibility with the historical and instrumental seismicity catalog adopted from the same work of \citet{Zare2014}. In addition, the dataset from IIEES also contained events reported with duration magnitude ($M_D$), as reported by the International Seismological Centre (ISC). In these cases we used the empirical relationship developed by \citet{Deniz2010} to convert $M_D$ to $M_w$. Although this relationship was derived based on a dataset for $M_w>=4.5$ events, we extrapolate the equation for lower magnitude and found that results to adjust well with the general trend, which should provide for good estimates of the $a$-value in the Gutenberg-Richter equation.

Last but not least important, we declustered the catalog prior to using it in the assessment of seismic hazard for the region of interest. This entailed the removal of foreshocks and aftershocks from the catalog to adjust the dataset to include only those events in agreement with a Poissonian model, which is a well-accepted model for seismic sources. Out of the original 7,478 earthquakes in the assembled catalog for the seismic regions of Azerbaijan, Alborz and Kopeh-Dagh, we removed 605 foreshocks and 2,917 aftershocks. The resulting final set of events in the catalog compiled for this study is shown in Fig.~\ref{fig:catalog}, which includes maps for both historical and instrumental seismicity.
